---
title: "p8105_mtp_yj2580"
author: "yj2580"
date: "10/20/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readxl)
library(tidyverse)
library(ggplot2)
library(gridExtra)
```

## Introduction
The study is targeted at the development of prominent exostosis young adults’ skulls emanating from the external occipital protuberance (EOP). Here authors try to demonstrate how a combination sex, the degree of forward head protraction and age can affect the presence of enlarged EOP (EEOP). To start with, they recruited 1200 people, aged 18–86.

## Problem 1
```{r}
data = 
  # load the data and skip first 8 extra rows
  read_excel("./data/p8105_mtp_data.xlsx", skip = 8) %>%
  # clean variables' names
  janitor::clean_names() %>%
  mutate(
  # fill in missing values with 0 for EOP size
    eop_size_mm = ifelse(is.na(eop_size_mm) == TRUE, "0", eop_size_mm),
    fhp_size_mm = ifelse(is.na(fhp_size_mm) == TRUE, "0", fhp_size_mm),
    eop_size_mm = as.numeric(eop_size_mm),
    fhp_size_mm = as.numeric(fhp_size_mm),
    # recode sex and age_group with illustrative characters
    sex = recode(sex, "0" = "female", "1" = "male"),
    age_group = recode(age_group, 
                       "1" = "<18", 
                       "2" = "18-30", 
                       "3" = "31-40", 
                       "4" = "41-50", 
                       "5" = "51-60", 
                       "6" = ">60", "7" = ">60", "8" = ">60"),
    # create ordered factors for categorical variables below
    sex = factor(sex, levels = c("female", "male")),
    age_group = factor(age_group, levels = c("<18", "18-30", "31-40", "41-50","51-60", ">60")),
    eop_size = factor(eop_size, levels =  c("0", "1", "2", "3", "4","5", "14.6")),
    eop_visibility_classification = factor(eop_visibility_classification, levels = c("0", "1", "2")),
    fhp_category = factor(fhp_category, levels = c("0", "1", "2", "3", "4", "5", "6", "7", "30.8"))
  )
#  In case of missing 14.6 in eop_size and 30.8 in fhp_category in subsequent analysis, I add them into levels.
```

There are `r nrow(data)` participants and `r ncol(data)` variables.
age_group = age groups according to decades
eop_size_mm = the size of the external occipital protuberance
eop_size = group eop_size_mm for every 5 mm
fhp_size_mm = the size of forward head protraction
fhp_category = group fhp_size_mm for every 10 mm

```{r}
data %>% 
  group_by(sex, age) %>%
  summarise(number_of_participants = n()) %>%
  knitr::kable()

data %>%
  group_by(age, sex) %>%
  summarise(number_of_participants = n()) %>%
  ggplot(aes(x=age, y=number_of_participants, fill = sex)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(x = "Age", y = "Number of participants", 
       title = "Age and gender distribution")
```

Sex distribution is almost even amongst age. Age distribution is nearly even from 18 to 72, although the amount is extremely high for group 61-year-old male and 51-year-old female. It is slightly right skewed in age distribution. (it will be even if grouped by age group, because ">60" group has a wider range.)

```{r}
data %>%
  janitor::clean_names() %>%
  mutate(
    age_correct = ifelse(age < 18, "<18",
                    ifelse(age <= 30, "18-30",
                      ifelse(age <= 40, "31-40",
                        ifelse(age <= 50, "41-50",  
                          ifelse(age <= 60, "51-60", ">60"))))),
    age_logical = ifelse(age_group == age_correct, "1", "0")
  ) %>%
  filter(age_logical == "0") %>%
  select(age, age_group, age_correct) %>%
  knitr::kable()

data %>%
  janitor::clean_names() %>%
  mutate(
    eop_size_mm = as.numeric(eop_size_mm),
    # correct misclassifications
    eop_size_correct = 
      ifelse(eop_size_mm<5, "0", 
          ifelse(eop_size_mm<10, "1",
              ifelse(eop_size_mm<15, "2", 
                  ifelse(eop_size_mm<20, "3", 
                        ifelse(eop_size_mm<25, "4", "5"))))),
    eop_size_logical = ifelse(eop_size == eop_size_correct, "1", "0")
  ) %>%
  filter(eop_size_logical == "0") %>%
  select(eop_size_mm, eop_size, eop_size_correct) %>%
  knitr::kable() 

data %>%
  janitor::clean_names() %>%
  mutate(
   eop_visibility_correct = 
      ifelse(eop_size_mm == 0, "0", 
             ifelse(eop_size_mm <= 5, "1", "2")),
   eop_visibility_logical = ifelse(eop_visibility_classification == eop_visibility_correct, "1", "0")
  ) %>%
  filter(eop_visibility_logical == "0") %>%
  select(eop_size_mm, eop_visibility_classification, eop_visibility_correct) %>%
  knitr::kable()

data %>%
  janitor::clean_names() %>%
  mutate(
   fhp_correct = 
      ifelse(fhp_size_mm <= 10, "0",
        ifelse(fhp_size_mm <= 20, "1",
          ifelse(fhp_size_mm <= 30, "2",
            ifelse(fhp_size_mm <= 40, "3",  
              ifelse(fhp_size_mm <= 50, "4",
                ifelse(fhp_size_mm <= 60, "5",
                  ifelse(fhp_size_mm <= 70, "6",
                    ifelse(fhp_size_mm <= 80, "7", "8" )))))))),
   fhp_logical = ifelse(fhp_category == fhp_correct, "1", "0")
  ) %>%
  filter(fhp_logical == "0") %>%
  select(fhp_size_mm, fhp_category, fhp_correct) %>%
  knitr::kable()
```

None of categorical variables correctly implements the definitions based on underlying continuous variables.

## Problem 2
```{r}
p1 = 
  data %>%
  ggplot(aes(x = age_group, y = fhp_size_mm, color = sex)) +
  geom_boxplot() +
  labs(x = "Age group", y = "Fhp size (mm)", title = "The distribution of fhp size")
 
eop_total = data %>%
  group_by(age_group, sex) %>%
  summarise(eop_total = n())
eop_enlarged = data %>%
  mutate(
    eop_logical = ifelse(eop_size == "0"|eop_size == "1", "normal", "enlarged")
  ) %>%
  filter(eop_logical == "enlarged") %>%
  group_by(age_group, sex) %>%
  summarise(eop_enlarged = n())
eop_rate = 
  left_join(eop_total, eop_enlarged, c("age_group", "sex")) %>%
  replace(is.na(.), 0) %>%
  filter(age_group != "<18") %>%
  mutate(eop_rate = eop_enlarged / eop_total)
p2 = 
  eop_rate %>%
  ggplot(aes(x = age_group, y = eop_rate, color = sex, group = sex)) + 
  geom_line() +
  geom_point() +
  labs(x = "Age group", y = "The rate of EEOP", title = "Trend of rate of EEOP among groups")

grid.arrange(p1, p2, ncol = 2, widths = c(1.5,1.5))
```

The fhp size distribution in every group is basically asymmetric, although it is slightly right skewed in group "51-60 female". Amongst the overall population, fhp size is right skewed and bell shaped with a peak around 25. 
EEOP rate is lower for female. For both gender, eop rate decreases from 18-30 to 41-50 and then increases from 41-50 to >60. Eop rate is obviously the highest for 18-30 age group.

```{r}
data %>%
  filter(age_group != "<18") %>%
  ggplot(aes(x = fhp_size_mm, y = eop_size_mm)) + 
  geom_point(alpha = .5) +
  geom_smooth(method = "lm") + 
  facet_grid(sex ~ age_group) +
  labs(x = "Fhp size", y = "Eop size", 
       title = "The association between FHP size and EOP size in each age and sex group")
```

In each group, fhp size is approximately positively associated with eop size, but the pattern is not obvious enough because dots are scattered randomly in each plot.  

## Problem 3

```{r}
# paper's target population is aged from 18 to 86
data_article = data %>%
  filter(age >= 18 & age <= 86)

# Author's statment: sample size is 1200
sample_size = data_article %>%
  nrow()
# Author's statment: 18–30 n=300, 31–40 n=200, 41–50 n=200, 51–60 n=200 and >60 n=300
data_article %>%
  group_by(age_group) %>%
  summarise(n = n()) %>%
  knitr::kable()
#statement :The mean FHP in the male cases examined was 28±15 mm, while that for the female cases was 24±11mm
data_article %>%
  group_by(sex) %>%
  summarise(mean = mean(fhp_size_mm), 
            stand_devi = sd(fhp_size_mm)) %>%
  knitr::kable()
# statment: the prevalence of EEOP to be 33% of the study population
eop_prev = data_article %>%
  mutate(
    eop_logical = ifelse(eop_size == "0"|eop_size == "1", "normal", "enlarged")
  ) %>%
  filter(eop_logical == "enlarged") %>%
  nrow()/sample_size
# statement: FHP >40 mm observed frequently (34.5%) in the over 60s cases
nfhp = data_article %>%
  filter(age_group == ">60") %>%
  nrow()
fhp_freq = data_article %>%
  filter(age_group == ">60", 
         fhp_size_mm > 40) %>%
  nrow()/nfhp
```

There are many inconsistencies. After calculation, sample size is `r sample_size`, not 1200. Sample sizes for each group are also inconsistent except for 51-60. Because age 45 is mis-classified into age group "<18", there is also one sample in "<18" which shouldn't exist. The mean FHP in the male was 28±15 mm, while for the female was 24±11mm, which is consistent. The prevalence of EEOP is `r eop_prev`, not 33%. The frequency of FHP>40mm is `r fhp_freq`, not 34.5%.

## Discussion
To sum up, age, sex and fhp size can all have some effect on EEOP. Female have less chance to get EEOP. Young adults are in greater danger of having EEOP, while middle-aged people are much safer. The association between fhp size and eop size is hard to clearly identify, maybe they are positively associated. Speaking of the flawness of this paper, authors made many mistakes during classifications which may have bad influence on subsequent analysis, leading to unconvincing conclusion. The statement "cell phones are causing horn growth" lacks adequate evidences. The positive association between fhp size and eop size is not strong enough at the first place. Furthermore, there are many causes of fhp other than cell phones. Also, causation can't be concluded unless they did a randomized control study. To address this hypothesis, I would collect information about the frequency of participants' cell phone use or how long they use cell phone everyday.

```{r}
wordcountaddin:::text_stats()
```

